# Activity Monitoring Data Analysis

## 0. Setup

```{r setup, echo=TRUE}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(gridExtra)
```

## 1. Read the dataset
```{r read, echo=TRUE}
MyData <- read.csv(file="activity.csv", header=TRUE, sep=",")
```

## 01. Histogram of the total number of steps taken each day

### Remove NA rows
```{r 02_01, echo=TRUE}
MyData2 <- MyData[complete.cases(MyData),]
```

### Calculate the total steps per day
```{r 02_02, echo=TRUE}
total_steps <- MyData2 %>%
     group_by(date) %>%
     summarise(avg_steps = mean(steps), sum_steps = sum(steps))
mean_steps <- mean(total_steps$sum_steps)
median_steps <- median(total_steps$sum_steps)
```

### Plot histogram chart
```{r 02_03, echo=TRUE}
ggplot(total_steps, aes(x = sum_steps)) + geom_histogram() + ggtitle("Sum of Steps per day")
```

## 3. Mean and median number of steps taken each day
The mean number of steps taken each day is `r mean_steps`.

The median number of steps taken each day is `r median_steps`.

## 4.Time series plot of the average number of steps taken
```{r 04_01, echo=TRUE}
ggplot(total_steps, aes(x = sum_steps)) + geom_histogram() + ggtitle("Sum of Steps per day - Original Data")
```

## 5.The 5-minute interval that, on average, contains the maximum number of steps

### Calculation
```{r 05_01, echo=TRUE}
steps_interval <- MyData2 %>%
     group_by(interval) %>%
     summarize(avg_steps = mean(steps))
max_steps <- max(steps_interval$avg_steps)
max_row <- which(steps_interval$avg_steps >= max_steps, arr.ind=T)
max_interval <- steps_interval[max_row, 1]

plot(steps_interval$interval, steps_interval$avg_steps, type="l", col="blue", xlab="5-minute interval", ylab="Average Steps per Day")
```

The 5-minute interval that, on average, contains the maximum number of steps is `r max_interval`.

## 6.Code to describe and show a strategy for imputing missing data

### Calculate the number of missing values
```{r 06_01, echo=TRUE}
missing_values <- sum(is.na(MyData$steps))
```
The number of missing values in the dataset is `r missing_values`.

### Complete missing values with the average for the interval
```{r 06_02, echo=TRUE}
missing_MyData <- MyData[!complete.cases(MyData),]
missing_MyData <- merge(missing_MyData[, c("steps", "date", "interval")], steps_interval[, c("interval", "avg_steps")])
missing_MyData$steps <- NULL
missing_MyData <- missing_MyData %>% rename (steps = avg_steps)
Complete_Data <- rbind(missing_MyData, MyData2)
```

## 7.Histogram of the total number of steps taken each day after missing values are imputed

### Calculate the total steps per day
```{r 07_01, echo=TRUE}
total_steps2 <- Complete_Data %>%
     group_by(date) %>%
     summarise(avg_steps = mean(steps), sum_steps = sum(steps))
mean_steps2 <- mean(total_steps2$sum_steps)
median_steps2 <- median(total_steps2$sum_steps)
```

### Plot histogram chart
```{r 07_02, echo=TRUE}
ggplot(total_steps2, aes(x = avg_steps)) + geom_histogram()+ ggtitle("Modified Data")
```

### Plot both histogram charts to compare how inputing missing values altered the results
```{r 07_03, echo=TRUE}
p1 <- ggplot(total_steps, aes(x = sum_steps)) + geom_histogram() + ggtitle("Original Data") + ylim(0, 15)
p2 <- ggplot(total_steps2, aes(x = sum_steps)) + geom_histogram()+ ggtitle("Modified Data") + ylim(0, 15)
grid.arrange(p1, p2, nrow = 1)
```

## 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

### Classify the days in Weekdays or Weekends
```{r 08_01, echo=TRUE}
Complete_Data$date <- as.POSIXct(Complete_Data$date)
Complete_Data <- Complete_Data %>% mutate(weekday = weekdays(date, abbreviate = TRUE))

i = 1
while (i <= nrow(Complete_Data)) {
     if (Complete_Data[i,"weekday"] == "Sat"  | Complete_Data[i,"weekday"] == "Sun") {
          Complete_Data[i,"weekday"] <- "weekend"
     } else {
          Complete_Data[i,"weekday"] <- "weekday"
     }
     i = i + 1
}
```

### Calculate the average of steps for each group 
```{r 08_02, echo=TRUE}
total_steps3 <- Complete_Data %>%
     group_by(interval, weekday) %>%
     summarise(avg_steps = mean(steps))
```

### Plot line chart for average steps
```{r 08_03, echo=TRUE}
ggplot(total_steps3, aes(x = interval, y = avg_steps, color = weekday)) + geom_line()
```